---
title: "vignette"
output: html_document
---

# Setup  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(dplyr)
library(janitor)
library(ggplot2)
library(plotly)
library(sf)
library(lubridate)
```

```{r yield}
yield <- read_sf("../data/yield/Keith Bowen-Colquitt-Shy North-Cotton.shp")

```

```{r boundary}
boundary <- read_sf("../data/boundary/Keith Bowen_Colquitt_Shy North.shp")
```

# Initial wrangling  
```{r yield_w}
yield_w <- yield %>%
  # Getting smaller sample for testing
  #sample_frac(.1) %>%
  clean_names() %>%
  # Setting lat/long and transforming to UTM
  st_transform(crs="EPSG:4326") %>%
  st_transform(crs=26916) %>%
  # Setting date, time, and speed
  mutate(datetime=ymd_hms(iso_time)) %>%
  mutate(date=date(datetime)) %>%
  arrange(datetime) %>%
  mutate(interval=interval(lag(datetime),datetime),
         intlenght_sec=int_length(interval)) %>%
  mutate(speed=distance/intlenght_sec) %>%
  # Transforming units
  mutate(distance_m=distance*0.3048,
         swath_m=swathwidth*0.3048,
         yield_lintkgha=(vryieldbal*480)*0.453592/0.404686,
         elev_m=elevation*0.3048,
         speed_kmh=speed*1.09728) %>%
  # Selecting what's needed
  dplyr::select(sectionid,
                heading,
                distance_m,
                swath_m,
                yield_lintkgha,
                elev_m,
                speed_kmh,
                datetime,
                date,
                geometry) %>%
  # Rounding
  mutate(across(heading:speed_kmh, ~round(.,2)))

yield_w
#rm(yield)

```

```{r boundary_w}
boundary_w <- boundary %>%
  clean_names() %>%
  # Setting lat/long and transforming to UTM
  st_transform(crs="EPSG:4326") %>%
  st_transform(crs=26916)
```


```{r bbb}
bbb <- boundary_w %>%
  st_bbox() 

bbb[2] <- bbb[2] + 80
bbb[4] <- bbb[4] + 80

bbb_w <- bbb %>%
  st_as_sfc() %>%
  st_buffer(-480)

yield_crop <- yield_w %>%
  st_crop(bbb_w)

ggplot()+
  #geom_sf(data=yield_crop, size=.5)+
  geom_sf(data=bbb_w, fill=NA)

```

```{r bitmap rotated}
rotang = -5

rot = function(a) matrix(c(cos(a), sin(a), -sin(a), cos(a)), 2, 2)

tran = function(geo, ang, center) (geo - center) * rot(ang * pi / 180) + center

center <- st_centroid(st_union(bbb_w))

bitmap <- st_make_grid(bbb_w, 
                       cellsize = .3,
                       crs = 26916)

bitmap_rot <- tran(bitmap, rotang, center) %>%
  st_set_crs(26916)


ggplot()+
  #geom_sf(data=bitmap, fill=NA, size=.01)+
  geom_sf(data=bitmap_rot, fill=NA, size=.1, color="purple")+
  geom_sf(data=bbb_w, fill=NA)+
  geom_sf(data=yield_crop, aes(color=heading), size=.1)+
  #scale_color_gradient(low="red", high="green")+
  coord_sf(xlim = c(713200.9+20, 713306.2-20), 
           ylim = c(3453737-1, 3453762+1), 
           expand = FALSE)+
  theme(legend.position = "none")

```

```{r a0k}
yield_crop %>%
  arrange(datetime) %>%
  mutate(a0k=swath_m*distance_m)
```

```{r overlap}
bitmap <- st_make_grid(boundary_w, cellsize = .3)

ggplot()+
  geom_sf(data=bitmap)+
  geom_sf(data=boundary_w, fill=NA, color="red")

```

# EDA  
```{r stats_preclean}
stats_preclean <- yield_w %>%
  as_tibble() %>%
  summarise(n=length(yield_lintkgha),
            min=min(yield_lintkgha, na.rm=T),
            median=median(yield_lintkgha, na.rm = T),
            mean=mean(yield_lintkgha, na.rm=T),
            max=max(yield_lintkgha, na.rm=T),
            sd=sd(yield_lintkgha, na.rm=T),
            cv=(sd/mean)*100
  ) %>%
  mutate(across(everything(), ~round(.,0))) %>%
  mutate(type="raw")

stats_preclean
```

```{r yield}
yield_w %>%
  ggplot(aes(x=yield_lintkgha))+
  geom_density()+
  labs(x="Lint yield (kg/ha)",
       y="Kernel density",
       title="Raw")+ 
  scale_x_continuous(labels = function(x) format(x, scientific = F))

ggsave(filename = "../output/1.rawdensity.png", width = 6, height = 4)

```

```{r distance}
yield_w %>%
  ggplot()+
  geom_sf(aes(color=distance_m), size=.5)+
  scale_color_gradient(low="red", high="green")

```

```{r distance v yield}
yield_w %>%
  ggplot(aes(x=distance_m, y=yield_lintkgha))+
  geom_point()
```
Maybe only distances >2 or 3 should be kept.  

```{r swath}
yield_w %>%
  dplyr::select(swath_m) %>%
  summary()
```
swathwidth is a constant value.  

```{r sectionid}
yield_w %>%
  mutate(sectionid=factor(sectionid)) %>%
  ggplot()+
  geom_sf(aes(color=sectionid), size=.5)

```
```{r heading}
yield_w %>%
  dplyr::select(heading) %>%
  summary

yield_w %>%
  mutate(headingf=case_when(
    heading > 0 & heading < 90 ~ "1",
    heading > 90 & heading < 180 ~ "2",
    heading > 180 & heading < 270 ~ "3",
    heading > 270 ~ "4"
  )) %>%
  #group_by(headingf) %>%
  #tally()
  ggplot()+
  geom_sf(aes(color=headingf), size=.5)
```

```{r timestamp}
yield_w %>%
  dplyr::select(datetime) %>%
  summary

```
harvest happened along 4 days.

```{r timestamp2}  
yield_w %>%
  mutate(date=factor(date)) %>%
  ggplot()+
  geom_sf(aes(color=date), size=.5)

yield_w %>%
  mutate(date=factor(date)) %>%
  filter(date=="2019-12-02") %>%
  ggplot()+
  geom_sf(aes(color=datetime), size=.5)+
  scale_color_datetime(low="red", high = "green")



```

```{r hours/d}
yield_w %>%
  as.data.frame() %>%
  mutate(datef=factor(date)) %>%
  group_by(datef) %>%
  dplyr::summarise(min=min(datetime),
                   max=max(datetime)) %>%
  mutate(durationday=interval(min, max),
         durationday_sec=int_length(durationday),
         durationday_h=durationday_sec/3600)

```
Worked ~24h on the 30th?  

```{r hours on the 30th}
yield_w %>%
  ggplot(aes(x=datetime, y=1))+
  geom_point()

```


```{r speed}
yield_w %>%
  dplyr::select(speed_kmh) %>%
  summary

```

# yieldeditR steps  
## min and max yield  
```{r min and max yield}
yield_ye1 <- yield_w %>%
  mutate(qhigh_pct=0.99,
         qlow_pct=0.1,
         scale=.25,
         absminy_kgha=10) %>%
  mutate(qhigh=quantile(yield_lintkgha, qhigh_pct),
         qlow=quantile(yield_lintkgha, qlow_pct),
         iqr=qhigh-qlow,
         miny=ifelse(qlow-(iqr*scale)<absminy_kgha,
                     absminy_kgha,
                     qlow-(iqr*scale)),
         maxy=qhigh+(iqr*scale)) %>%
  mutate(flag=case_when(
    yield_lintkgha < miny ~ "miny",
    yield_lintkgha > maxy ~ "maxy",
    T ~ NA_character_
  )) %>%
  dplyr::select(-qhigh_pct, -qlow_pct, -scale,
                -qhigh, -qlow, -iqr)

yield_ye1 %>%
  as.data.frame() %>%
  group_by(flag) %>%
  tally() #%>%
summarise(N=sum(n))

```
```{r stats_minmaxyield}
stats_minmaxyield <- yield_ye1 %>%
  as_tibble() %>%
  filter(!grepl("miny|maxy", flag)) %>%
  summarise(n=length(yield_lintkgha),
            min=min(yield_lintkgha, na.rm=T),
            median=median(yield_lintkgha, na.rm = T),
            mean=mean(yield_lintkgha, na.rm=T),
            max=max(yield_lintkgha, na.rm=T),
            sd=sd(yield_lintkgha, na.rm=T),
            cv=(sd/mean)*100
  ) %>%
  mutate(across(everything(), ~round(.,0))) %>%
  mutate(type="minmaxyield")

stats_minmaxyield 

```

```{r yield_ye1 density}
yield_ye1 %>%
  as_tibble() %>%
  filter(!grepl("miny|maxy", flag)) %>%
  ggplot(aes(x=yield_lintkgha))+
  geom_density(fill="forestgreen", alpha=.5)
```

## min and max speed  
```{r min and max speed}
yield_ye2 <- yield_ye1 %>%
  mutate(qhigh_pct=0.9,
         qlow_pct=0.08,
         scale=.18,
         absminy_kmh=0.8) %>%
  mutate(qhigh=quantile(speed_kmh, qhigh_pct, na.rm=T),
         qlow=quantile(speed_kmh, qlow_pct, na.rm=T),
         iqr=qhigh-qlow,
         minv=ifelse(qlow-(iqr*scale)<absminy_kmh,
                     absminy_kmh,
                     qlow-(iqr*scale)),
         maxv=qhigh+(iqr*scale)) %>%
  mutate(flag=case_when(
    speed_kmh < minv & is.na(flag) ~ "minv",
    speed_kmh < minv & !is.na(flag) ~ paste0(flag,"-minv"),
    speed_kmh > maxv & is.na(flag) ~ "maxv",
    speed_kmh > maxv & !is.na(flag) ~ paste0(flag,"-maxv"),
    !is.na(flag) ~ flag,
    T ~ NA_character_
  )) %>%
  dplyr::select(-qhigh_pct, -qlow_pct, -scale,
                -qhigh, -qlow, -iqr)

yield_ye2 %>%
  as.data.frame() %>%
  group_by(flag) %>%
  tally() #%>%
summarise(N=sum(n))

```

```{r stats_minmaxsp}
stats_minmaxsp <- yield_ye2 %>%
  as_tibble() %>%
  filter(!grepl("minv|maxv", flag)) %>%
  summarise(n=length(yield_lintkgha),
            min=min(yield_lintkgha, na.rm=T),
            median=median(yield_lintkgha, na.rm = T),
            mean=mean(yield_lintkgha, na.rm=T),
            max=max(yield_lintkgha, na.rm=T),
            sd=sd(yield_lintkgha, na.rm=T),
            cv=(sd/mean)*100
  ) %>%
  mutate(across(everything(), ~round(.,0))) %>%
  mutate(type="minmaxspeed")

stats_minmaxsp 

```

```{r yield_ye2 density}
yield_ye2 %>%
  as_tibble() %>%
  filter(!grepl("minv|maxv", flag)) %>%
  ggplot(aes(x=yield_lintkgha))+
  geom_density(fill="purple", alpha=.5)
```
```{r stats_minmaxyieldsp}
stats_minmaxyieldsp <- yield_ye2 %>%
  as_tibble() %>%
  filter(!grepl("miny|maxy|minv|maxv", flag)) %>%
  summarise(n=length(yield_lintkgha),
            min=min(yield_lintkgha, na.rm=T),
            median=median(yield_lintkgha, na.rm = T),
            mean=mean(yield_lintkgha, na.rm=T),
            max=max(yield_lintkgha, na.rm=T),
            sd=sd(yield_lintkgha, na.rm=T),
            cv=(sd/mean)*100
  ) %>%
  mutate(across(everything(), ~round(.,0))) %>%
  mutate(type="minmaxyield+minmaxspeed")

stats_minmaxyieldsp 

```

```{r yield_ye1 and 2 plot}
yield_ye2 %>%
  as_tibble() %>%
  filter(!grepl("miny|maxy|minv|maxv", flag)) %>%
  ggplot(aes(x=yield_lintkgha))+
  geom_density(fill="red", alpha=.5)

```

## pos  

For the POS parameter, the original paper uses the IQR of easting and northing + 40 m. I think a negative buffer from the field boundaries would work just as well.
```{r pos}
yield_ye3 <- yield_ye2 %>%
  st_coordinates() %>%
  as_tibble() %>%
  dplyr::select(-Z) %>%
  bind_cols(yield_ye2) %>%
  mutate(qhigh_pct=0.995,
         qlow_pct=0.005,
         buffer_m=40) %>%
  mutate(qhigh_x=quantile(X, qhigh_pct, na.rm=T),
         qlow_x=quantile(X, qlow_pct, na.rm=T),
         iqr_x=qhigh_x-qlow_x,
         minx=qlow_x-buffer_m,
         maxx=qhigh_x+buffer_m) %>%
  mutate(qhigh_y=quantile(Y, qhigh_pct, na.rm=T),
         qlow_y=quantile(Y, qlow_pct, na.rm=T),
         iqr_y=qhigh_y-qlow_y,
         miny=qlow_y-buffer_m,
         maxy=qhigh_y+buffer_m) %>%
  mutate(flag=case_when(
    X < minx  | X > maxx |
      Y < miny | Y > maxy & is.na(flag) ~ "pos",
    X < minx | X > maxx |
      Y < miny | Y > maxy & !is.na(flag) ~ paste0(flag,"-pos"),
    !is.na(flag) ~ flag,
    T ~ NA_character_
  )) %>%
  dplyr::select(-qhigh_pct, -qlow_pct, -buffer_m,
                -qhigh_x, -qlow_x, -iqr_x,
                -qhigh_y, -qlow_y, -iqr_y)

yield_ye3 %>%
  as.data.frame() %>%
  group_by(flag) %>%
  tally
```
Nothing was flagged.
Perhaps using the boundary would be better

```{r pos with boundary}
yield_ye3 <- yield_ye2 %>%
  st_within(st_buffer(boundary_w, -20), sparse = F) %>%
  as_tibble() %>%
  rename(pos=V1) %>%
  bind_cols(yield_ye2) %>%
  st_as_sf() %>%
  mutate(flag=case_when(
    pos == F & is.na(flag) ~ "pos",
    pos == F & !is.na(flag) ~ paste0(flag,"-pos"),
    !is.na(flag) ~ flag,
    T ~ NA_character_
  )) 


yield_ye3 %>%
  group_by(pos) %>%
  tally()

yield_ye3 %>%
  group_by(flag) %>%
  tally()


ggplot()+
  geom_sf(data=yield_ye3, size=.5, aes(color=pos))+
  geom_sf(data=boundary_w, fill=NA, color="red")+
  geom_sf(data=st_buffer(boundary_w,-20), fill=NA, color="blue")

```
```{r stats_pos}
stats_pos <- yield_ye3 %>%
  as_tibble() %>%
  filter(!grepl("pos", flag)) %>%
  summarise(n=length(yield_lintkgha),
            min=min(yield_lintkgha, na.rm=T),
            median=median(yield_lintkgha, na.rm = T),
            mean=mean(yield_lintkgha, na.rm=T),
            max=max(yield_lintkgha, na.rm=T),
            sd=sd(yield_lintkgha, na.rm=T),
            cv=(sd/mean)*100
  ) %>%
  mutate(across(everything(), ~round(.,0))) %>%
  mutate(type="pos")

stats_pos 

```

```{r yield_ye3 plot}
yield_ye3 %>%
  as_tibble() %>%
  filter(!grepl("pos", flag)) %>%
  ggplot(aes(x=yield_lintkgha))+
  geom_density(fill="pink", alpha=.5)

```

```{r stats_minmaxyieldsppos}
stats_minmaxyieldsppos <- yield_ye3 %>%
  as_tibble() %>%
  filter(!grepl("miny|maxy|minv|maxv|pos", flag)) %>%
  summarise(n=length(yield_lintkgha),
            min=min(yield_lintkgha, na.rm=T),
            median=median(yield_lintkgha, na.rm = T),
            mean=mean(yield_lintkgha, na.rm=T),
            max=max(yield_lintkgha, na.rm=T),
            sd=sd(yield_lintkgha, na.rm=T),
            cv=(sd/mean)*100
  ) %>%
  mutate(across(everything(), ~round(.,0))) %>%
  mutate(type="minmaxyield+minmaxspeed+pos")

stats_minmaxyieldsppos 

```

```{r yield_ye1 2 and 3 plot}
yield_ye3 %>%
  as_tibble() %>%
  filter(!grepl("miny|maxy|minv|maxv|pos", flag)) %>%
  ggplot(aes(x=yield_lintkgha))+
  geom_density(fill="yellow", alpha=.5)

```


## overlap  
```{r overlap}
bitmap <- st_make_grid(boundary_w, cellsize = .3)

ggplot()+
  geom_sf(data=bitmap)+
  geom_sf(data=boundary_w, fill=NA, color="red")

```
# Report 
```{r yield}
yield_w %>%
  ggplot()+
  geom_sf(aes(color=yield_lintkgha), size=.2)+
  scale_color_gradient(low="red", high="green")+
  labs(color="Lint yield\n(kg/ha)")

```

```{r yield}
yieldmap_plot <- yield_w %>%
  filter(yield_lintkgha<5000) %>%
  ggplot()+
  geom_sf(aes(color=yield_lintkgha), size=.2)+
  scale_color_gradient(low="red", high="green")+
  labs(color="Lint yield\n(kg/ha)", title="Raw")+
  theme(legend.position = "bottom",
        legend.key.width= unit(1.5, 'cm'))

yieldmap_plot

ggsave(yieldmap_plot, filename =  "../output/2.rawmap.png", width = 7, height = 6)

```

## Table  
```{r table individual filters}
stats_preclean %>%
  bind_rows(stats_minmaxyield) %>%
  bind_rows(stats_minmaxsp) %>%
  bind_rows(stats_pos) %>%
  mutate(n_pctoriginal=round(n/142944,2)) %>%
  dplyr::select(type, n, n_pctoriginal, everything()) %>%
  readr::write_csv("../output/table1.csv")

```

```{r table combined filters}
stats_comb <- stats_preclean %>%
  bind_rows(stats_minmaxyield) %>%
  bind_rows(stats_minmaxyieldsp) %>%
  bind_rows(stats_minmaxyieldsppos) %>%
  mutate(n_pctoriginal=round(n/142944,2)) %>%
  dplyr::select(type, n, n_pctoriginal, everything())

stats_comb

readr::write_csv(stats_comb, "../output/table2.csv")
```

## Plot  
```{r raw yield density}
yield_w %>%
  ggplot(aes(x=yield_lintkgha))+
  geom_density()+
  labs(y="Kernel density", x="Lint yield (kg/ha)")

```

```{r density plot combined filters}
yield_w %>%
  #filter(yield_lintkgha<5000) %>%
  mutate(type="raw") %>%
  bind_rows(yield_ye1 %>%
              as_tibble() %>%
              filter(!grepl("miny|maxy", flag)) %>%
              mutate(type="minmaxyield")) %>%
  bind_rows(yield_ye2 %>%
              as_tibble() %>%
              filter(!grepl("miny|maxy|minv|maxv", flag)) %>%
              mutate(type="minmaxyield+minmaxspeed")) %>%
  bind_rows(yield_ye3 %>%
              as_tibble() %>%
              filter(!grepl("miny|maxy|minv|maxv|pos", flag)) %>%
              mutate(type="minmaxyield+minmaxspeed+pos")) %>%
  ggplot()+
  #ggridges::geom_density_ridges
  geom_density(aes(x=yield_lintkgha, 
                   #y=type, 
                   fill=type), 
               alpha=.6)+
  #geom_vline(data=stats_comb, aes(xintercept=mean))+
  theme(legend.position = "bottom")+
  labs(x="Lint yield (kg/ha)", y="Kernel density")




```

```{r map combined}
yield_ye3 %>%
  filter(!is.na(flag)) %>%
  ggplot()+
  geom_sf(aes(color=flag), size=.2)+
  guides(color = guide_legend(override.aes = list(size = 3)))
  

ggsave(filename =  "../output/3.removedmap.png", width = 7, height = 6)

```
```{r}
yield_ye3 %>%
  as_tibble() %>%
  group_by(flag) %>%
  tally() %>%
  arrange(desc(n)) %>%
  readr::write_csv("../output/table3.csv")
```


```{r yield map filtered}
yieldmapclean_plot <- yield_ye3 %>%
  filter(is.na(flag)) %>%
  ggplot()+
  geom_sf(aes(color=yield_lintkgha), size=.2)+
  scale_color_gradient(low="red", high="green")+
  labs(color="Lint yield\n(kg/ha)", title="Cleaned")+
  theme(legend.position = "bottom",
        legend.key.width= unit(1.5, 'cm'))

yieldmapclean_plot



```

```{r pre-post cleaning map}
library(patchwork)
yieldmap_plot+yieldmapclean_plot

ggsave(filename =  "../output/4.raw-cleanmap.png", width = 12, height = 6)
```


# TO-DO  
Need to find out:  
- unit on distance: it is feet  
- unit on swath: probably feet    
- unit on wet mass: lbs
- how is yield calculated (swath, distance, wet mass?)  
- is there a moisture column?  check on Ops center. checked, didn't find  
- Has delay already being fixed on ops? If not, what would be a good estimate? 2 secs?

Need to fix projection  (DONE)

# Needed for yield editor:  
- Yield  
- Speed  

# Notes:  
Cotton delay is a lot less than grain delay (2 secs vs. 12 secs).  

# Questions to JD  
- how is yield calculated (swath, distance, wet mass? moisture?)  
- is there a moisture column?  
- Has delay already being fixed on ops? If not, what would be a good estimate? 2 secs?
- Have any other transformations been made to the data on ops?  





